Concurrent Programming with CXL

### 2020-06-23 book/paper.tex
Latest

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RWfair.cxl}
\end{code}
\caption{[RWfair.cxl] Reader/Writer Lock SBS implementation addressing fairness.}
\label{fig:RWfair}
\end{figure}

@@

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RWqueue.cxl}
\end{code}
\caption{[RWqueue.cxl] Reader/Writer Lock SBS implementation using a queue.}
\label{fig:RWqueue}
\end{figure}

@@

\texttt{wait} releases the mutex and blocks while trying
to procure the condition variable (which is a semaphore that is 0 when
\texttt{wait} is invoked).
\texttt{signal} vacates the condition variable, passing the baton to
a process trying to procure it, and then procures the mutex.
The use of \texttt{wait} and \texttt{signal} ensures that \texttt{P} and
\texttt{V} operations alternate as prescribed by the SBS technique.
Our implementations of reader/writer locks can be easily changed to
use \texttt{wait} and \texttt{signal} instead of using \texttt{P}
\texttt{V}.

@@

As in Hoare monitors, there is a hidden mutex associated with each monitor,
and the mutex is automatically acquired upon entry to a method and released
upon exit.
Mesa monitors also have condition variable that a process can wait on.
Like in Hoare monitors, the \texttt{wait} operation release the mutex.
The most important difference is in what \texttt{signal} does.
To make the distinction more clear, we shall call the corresponding Mesa
operation \texttt{notify} rather than \texttt{signal}.
\index{notify}
When a process $p$ invokes \texttt{notify}, it does not immediately pass 
the baton to some process $q$ requiring the baton.  Instead, $p$ holds onto
the baton until it leaves the monitor.  At that point $q$ gets the baton,
assuming there even was such a process $q$.

@@

Figure~\ref{fig:monitors} illustrates the difference with a drawing.
Here a bathroom represents the critical section, allowing only one
process at a time say.  The bedrooms represent condition variables.
There is some condition associated with each bedroom.
In the hall are processes waiting to enter the critical section.

@@

\section*{Exercises}
\begin{problems}
\item Figure~\ref{fig:bank} shows an implementation of a bank with various
accounts and transfers between those accounts.
Unfortunately, running the test reveals that it sometimes leaves unterminated
processes.  Can you fix the problem?
\end{problems}

>>>>

\begin{figure}
\begin{code}
{\small
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RWfair.cxl}
}
\end{code}
\caption{[RWfair.cxl] Reader/Writer Lock SBS implementation addressing fairness.}
\label{fig:RWfair}
\end{figure}

@@

\begin{figure}
\begin{code}
{\small
\VerbatimInput[xleftmargin=5mm,numbers=left]{code/RWqueue.cxl}
}
\end{code}
\caption{[RWqueue.cxl] Reader/Writer Lock SBS implementation using a queue.}
\label{fig:RWqueue}
\end{figure}

@@

\texttt{wait} releases the mutex and blocks while trying
to procure the condition variable (which is a semaphore that is 0 when
\texttt{wait} is invoked).
\texttt{signal} vacates the condition variable, passing the baton to
a process trying to procure it, and then procures the mutex.
The use of \texttt{wait} and \texttt{signal} ensures that \texttt{P} and
\texttt{V} operations alternate as prescribed by the SBS technique.
Our implementations of reader/writer locks can be easily changed to
use \texttt{wait} and \texttt{signal} instead of using \texttt{P}
and \texttt{V}.

@@

As in Hoare monitors, there is a hidden mutex associated with each monitor,
and the mutex is automatically acquired upon entry to a method and released
upon exit.
Mesa monitors also have condition variable that a process can wait on.
Like in Hoare monitors, the \texttt{wait} operation releases the mutex.
The most important difference is in what \texttt{signal} does.
To make the distinction more clear, we shall call the corresponding Mesa
operation \texttt{notify} rather than \texttt{signal}.
\index{notify}
When a process $p$ invokes \texttt{notify}, it does not immediately pass 
the baton to some process $q$ requiring the baton (assuming there even
is such a process).  Instead, $p$ holds onto
the baton until it leaves the monitor.  At that point, any process that
was notified will have a chance to obtain the baton.

@@

Figure~\ref{fig:monitors} illustrates the difference with a drawing.
Here a bathroom represents the critical section, allowing only one
process at a time.  The bedrooms represent condition variables.
There is some condition associated with each bedroom.
In the hall are processes waiting to enter the critical section.

@@

\section*{Exercises}
\begin{problems}
\item Figure~\ref{fig:bank} shows an implementation of a bank with various
accounts and transfers between those accounts.
Unfortunately, running the test reveals that it sometimes leaves unterminated
processes.  Can you fix the problem?
\item Add a method \texttt{total()} to the solution of the previous question
that computes the total over all balances.
It needs to obtain a lock on all accounts.  Make sure that
it cannot cause deadlock.
\item Use method \texttt{total()} to ensure that the sum total of the bank
accounts at the beginning of the test is the same as at the end.
\end{problems}
